AppPAL for Android

### 2015-08-05 paper.tex
updates and editing

% \pdfpagewidth=8.5truein{}
% \pdfpageheight=11truein{}

\usepackage{graphicx}
% \renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
% \renewcommand{\floatpagefraction}{0.99}
% \pagestyle{plain} % page numbers

\usepackage{defaultstyle}
\usepackage{localstyle}

@@

\begin{abstract}
  % \begin{itemize}
  %   \item \comment{Sentence 1: State the problem}
  Users must judge apps by the information shown to them by the store.
  %   \item \comment{Sentence 2: Why is it an interesting problem}
  Employers rely on employees enforcing company policies correctly on devices at their workplace.
  Some users take time to pick apps with care whilst others do not.
  Users feel frustrated when they realise what data the apps have access to.
  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
  %   \item \comment{Sentence 3: What does my solution achieve}
  We present AppPAL: a policy language to enforce policies about apps.
  AppPAL Policies use statements from third parties and delegation relationships to
  %   \item \comment{Sentence 4: What follows from my solution}
  give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
  % \end{itemize}
\end{abstract}

@@

%\comment{The problem is\ldots}
Finding the right apps can be tricky:
  users need to work out which apps are well written, which are not going to abuse their data, which ones will work in the way they want
  and to find the apps which suit how they want to use their device.
This can be difficult as it isn't obvious how apps use the data each has access to.

App stores give some information about their apps; such as permissions, descriptions of the app and screenshots as well as review scores.
Android apps currently show a confusing list of permissions when they're first installed.
Soon new Android apps will display permissions requests when the app first tries to access sensitive data such as contacts or location information.
Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the user.

Not all apps are suitable.
A large amount of \ac{pus} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (bring your own device or BYOD).
An employer may wish to restrict which apps their employees can use.
The IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They may apply their own personal security policy when downloading and running apps.

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.3\linewidth]{figures/facebook.png}
%   \caption{Security information presented when a user downloads the Facebook app.}
% \label{fig:facebook}
% \end{figure}

%\comment{My contribution is\ldots}
We present AppPAL, an authorization logic for reasoning about apps.
The language is an instantiation of SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.

Using AppPAL we can say an app is installable:
\begin{lstlisting}
"Alice" says "com.rovio.angrybirds" isInstallable.
\end{lstlisting}
We can delegate decisions to others:
\begin{lstlisting}
"Alice" says "Bob" can-say 0 App isCool.
\end{lstlisting}
We can run analysis tools, and solve constraints:
\begin{lstlisting}
"Claire" says App isMalware
  if App isAnApp
  where
    virusScanner(App) = true.
\end{lstlisting}
Constraints allow up to express facts that are true at some times but not others:
\begin{lstlisting}
"Emma" says "com.facebook.katana" isRunnable
  where timeOfDay() > 1700.
\end{lstlisting}

Specifically we have:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem}.)
    We show how the trust relationships in a system can be difficult to see;
      and how AppPAL makes the trust between users explicit.
  \item
    Shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.
    We have described what AppPAL would check for an app to meet this policy and how these statements might be collected.

  \item Implemented the AppPAL language on Android and the JVM.
    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}),
      how we evaluate it (\autoref{ssec:eval}),
      and some idiomatic policies (\autoref{ssec:idioms}).

  \item When things go wrong we also need to attribute blame.
    Authorization logic can help make trust relationships explicit.
    We show a recent example of a broken trust relationship (\autoref{ssec:language})
    and how our authorization logic makes explicit where the breach of trust is.

\end{itemize}

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone though---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
could configure Alice's device to disallow apps from outside the Google Play Store (and soon in the newest version of Android let Emma set the permissions of each app on an app by app basis).
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are ever leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from stores which ensure the only apps they see are the apps which meet their policies.

@@

% \begin{figure}
%   \includegraphics{figures/bob-policy.eps}
%   \caption{Pictorial representation of Alice's app install policy.}
% \label{fig:emmas_policy}
% \end{figure}

@@

"emma" says "google-play" can-say 0
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=14]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
    This helps keep the number of  assertions in the policy low aiding readability.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself.

@@

\section{AppPAL}
\label{sec:details}

AppPAL is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.

\subsection{The Language}
\label{ssec:language}

@@

In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of predicates, and constraints for their SecPAL open.
With AppPAL we make explicit our predicates and how they relate to Android.
Some of these predicates require arguments.
We add a light typing system to help ensure these predicates have the correct arguments.

Specifically we define the following predicates:
\begin{description}
  \item[\texttt{App isRunnable}]
    Used to indicate that an App meets the install policy for the device.
    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.

  \item[\texttt{App isAnApp}, \texttt{Policy isAPolicy}\ldots]
    Expresses a simple typing relations.
    AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
    These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
    This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.

  \item[\texttt{Policy isMetBy(App)}]
    Policies let you split app behaviour into sub-policies.
    For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
    Splitting policies allows greater control of how each one is checked.
    We can delegate checking a policy to an expert using the \code{can-say} statement.
    We can specify more detailed checks using the \code{can-act-as} statements.

  \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
    A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
\end{description}

@@

Consider the \emph{no-tracking-policy} from earlier.
In Bob's policy we checked this by checking the permissions of the app.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
Our policy now becomes:

@@

AppPAL also helps attribute blame when things go wrong.
By modelling the trust relationships in systems we can work out precisely where mistakes were made.
A recent example demonstrating this is \textsc{CVE-2015-2077}.
Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
Unfortunately they had configured it to ignore Superfish.
The malware was an ad framework designed to show users products similar to those they viewed on the web.
Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.

From a user's perspective there was a delegation of trust to Lenovo to detect malware.
Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
A user might wonder where the breach of trust occurred and who is to blame?
Is it the \ac{av} failing to spot the malware?
Has someone else configured the software incorrectly?
How should they fix the problem?

If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
\begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
"user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
"lenovo" says "mcafee" can-say inf File isSafe.
"lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
\end{lstlisting}
The fault lies with \autoref{lst:lenovo_unsafe}.
Lenovo has caused the problem (they \emph{said} it was safe).
The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.

\subsection{Evaluation}
\label{ssec:eval}

To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
We do not use the DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm suggested by Becker~\etal.
Rather we implement the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.
Like Becker~\etal we make use of an assertion context to store known statements.
We also store intermediate results to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.

\begin{figure}
\[\footnotesize
% AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
% ----------------------------------------------------------- (can-act-as)
%                 AC, D ⊧ A says B verbphrase
\infer[\text{\tiny can-act-as}]{
  AC, D \models \text{\tt A says B verbphrase}}{%
  AC, D \models \text{\tt A says B can-act-as C} & %
  AC, D \models \text{\tt A says C verbphrase}}
\]
\[\footnotesize
% AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
% ----------------------------------------------------- (can-say)
%                   AC, ∞ ⊧ A says fact
\infer[\text{\tiny can-say}]{
  AC, \infty \models \text{\tt A says fact}}{%
  AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
  AC, D \models \text{\tt B says fact}}
\]
\[\footnotesize
% (A says fact if fact₁, …, factₖ, c) ∈ AC
% AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
% ----------------------------------------------------------- (cond)
%                    AC, D ⊧ A says factθ
\infer[\text{\tiny cond}]{
  AC, D \models \text{\tt A says fact$\theta$}}{
  \begin{array}{c}
    (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
    \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
  \end{array} &
  \models \mathtt{c}\theta &
  \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
\]
\caption{AppPAL's evaluation rules.}
\label{fig:rules}
\end{figure}

@@

In SecPAL there is no type.
All entities are either constants (implemented as strings) or variables.
There is no differentiation made between a \emph{voiced} constant (who says assertions),
and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
In AppPAL the subjects are usually apps or policies.
Apps do not utter facts;
  so there is no need to check them as potential speakers of \emph{can-say} statements.
Similarly if a constant never appears as a subject of a statement, as most speakers do not,
  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
This means we need to search less constants when attempting to replace a variable, speeding evaluation.

\begin{figure}
\begin{minipage}[b]{0.5\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
# ac, d |-? q (using previous-results table rt)
def evaluate(ac, rt, q, d)
  return rt[q, d] if rt.contains q, d
  p = cond(ac, rt, q, d)
  return (Proven, rt.update q, d, p) if p.isValid
  p = canSay_CanActAs(ac, rt, q, d)
  return (Proven, rt.update q, d, p) if p.isValid
  return (Failure, rt.update q, d, Failure)

# Try can-say and can-act-as
def canSay_CanActAs(ac, rt, q, d)
  ac.constants.each do |c|
    if c.is_a :subject
      p = canActAs ac, rt, q, d
      return Proven if p.isValid
    elsif c.is_a :speaker
      p = canSay ac, rt, q d
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\begin{minipage}[b]{0.5\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily\scriptsize, keywordstyle=\scriptsize\slshape, columns=flexible]
# Try cond rule
def cond(ac, rt, q, d)
  ac.add q.fetch if q.isFetchable
  ac.assertions.each do |a|
    if (u = q.unify a.consequent) &&
       (a = u.substitution a).variables == none
      return checkConditions ac, rt, a, d
  return Failure

# Evaluate the antecedents
def checkConditions(ac, rt, a, d)
  getVarSubstitutions(a,ac.constants).each do |s|
    sa = s.substitute a
    if sa.antecedents.all
        { |a| evaluate(ac, rt, a, d).isValid }
      p = evaluateC sa.constraint
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\caption{Partial-pseudocode for AppPAL evaluation.}
\label{fig:pseudocode}
\end{figure}

\subsection{Policy Examples}
\label{ssec:idioms}

When writing policies some schemes often come up.
We give examples of policies and queries and show how they can be implemented in AppPAL.

>>>>

\usepackage{graphicx}
\usepackage{defaultstyle}
\usepackage{localstyle}

@@

\begin{abstract}
  Users must judge apps by the information shown to them by the store.
  Employers rely on employees enforcing company policies correctly on devices at their workplace.
  Some users take time to pick apps with care whilst others do not.
  Users feel frustrated when they realise what data the apps have access to.
  They want greater control over what data they give away but they do not want to spend time reading permissions lists.
  We present AppPAL: a policy language to enforce policies about apps.
  AppPAL Policies use statements from third parties and delegation relationships to
  give us a rigorous and flexible framework for enforcing and comparing and app policies; and the trust relationships surrounding them.
  % \end{itemize}
\end{abstract}

@@

Finding the right apps can be tricky.
Users need to work out which apps are well written, which are not going to abuse their data, which will work in the way they want
  and to find the apps which suit how they want to use their device.
This can be difficult as it isn't obvious how apps use the data each has access to.

App stores give some information about their apps; descriptions of the app and screenshots as well as review scores.
Android apps show a list of permissions when they're first installed.
Soon new Android apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not all understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the user.

Not all apps are suitable.
A large amount of \ac{pup} is being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work (\ac{byod}).
An employer may wish to restrict which apps their employees can use.
The IT department may set a policy to prevent information leaks.
Some users worry apps will misuse their personal data;
  such a user avoids apps which can access their location, or address book.
They may apply their own personal security policy when downloading and running apps.

@@

We present AppPAL, an authorization logic for reasoning about apps.
The language is an instantiation of SecPAL~\cite{Becker:2006vh} with constraints and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
The implementation allows us to enforce the policies on a device.
We can express trust relationships amongst these parties; use constraints to do additional checks.
This lets us enforce deeper and more complex policies than existing tools such as Kirin~\cite{Enck:2009ko}.

Using AppPAL we write policies for work and home, and decide which policy to use using a user's location, or the time of day:
\begin{lstlisting}
"alice" says App isRunnable
  if "home-policy" isMetBy(App)
  where At("work") = false.

"alice" says App isRunnable
  if "work-policy" isMetBy(App)
  where BeforeHourOfDay("17") = true.
\end{lstlisting}
We can delegate policy specification to third parties or roles, and assign principals to those rolls:
\begin{lstlisting}
"alice" says "it-department" can-say 0 "work-policy" isMetBy(App).
"alice" says "bob" can-act-as "it-department".
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization:
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissions-policy" isMetBy(App).
"alice" says "permissions-policy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

% Using AppPAL we can say an app is installable:
% \begin{lstlisting}
% "Alice" says "com.rovio.angrybirds" isInstallable.
% \end{lstlisting}
% We can delegate decisions to others:
% \begin{lstlisting}
% "Alice" says "ITDepartment" can-say 0 App isInstallable.
% \end{lstlisting}
% We can run analysis tools, and solve constraints:
% \begin{lstlisting}
% "Claire" says App isMalware
%   if App isAnApp
%   where
%     virusScanner(App) = true.
% \end{lstlisting}
% Constraints allow up to express facts that are true at some times but not others:
% \begin{lstlisting}
% "Emma" says "com.facebook.katana" isInstallable
%   where timeOfDay() > 1700.
% \end{lstlisting}

\noindent
Specifically our contribution is:
\begin{itemize}
  \item
    Described a scenario where an employer has a policy they want to enforce for their employees~(\autoref{sec:problem});
    and shown how the employer's policy could be implemented using AppPAL~(\autoref{sec:idea}) and installed on Alice's phone.

  \item Implemented the AppPAL language on Android and the JVM.
    We show how the language can describe properties of Android apps and Android security poilicies (\autoref{ssec:language}) and how we can check policies (\autoref{ssec:eval}).

  \item Shown the need for policy tools by demonstrating the privacy paradox holds for user privacy policies with app installations (\autoref{sec:demonstation}).
\end{itemize}

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone though---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}} could configure Alice's device to disallow apps from outside the Google Play Store (and in the newest version of Android let Emma set the permissions of each app on an app by app basis~\cite{AndroidMPermission:2015uq}).
AppPAL is designed to build on these tools, but make the trust and delegation relationships explicit and formalize the decision making process.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\cite{Enck:2009ko}, Flowdroid~\cite{Fritz:2013vi} or DroidSafe~\cite{Gordon:2015et} could check that the locations are ever leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

We propose a change to the ecosystem, shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from stores which ensure the only apps they see are the apps which meet their policies.
Policies can be combined

@@

@@

"emma" says "google-play" can-say 0
  App isReputable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15]
"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
    This helps keep the number of  assertions in the policy low aiding readability.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if running on device.

@@

\section{AppPAL}
\label{sec:details}
\label{ssec:language}

AppPAL is an instantiation of SecPAL for Android app installation policies.
It is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.

@@

In the Becker~\etal's paper~\cite{Becker:2006vh} they leave the choice of
predicates, and constraints for their SecPAL open.  With AppPAL we make
explicit our predicates and how they relate to Android.  Some of these
predicates require arguments.  AppPAL policies typically make use of the
predicates and constraints in \autoref{tab:predicates}.  Additional predicates
can be created in the policy files, but adding or modifying constraints
requires a code change\footnote{But a small one.  The Android version of the
\texttt{hasPermission} constraint, for example, uses the Android package
manager to determine what permissions an app requests, but the Java version
uses the Android platform tools.  This required a code change to one function.}

\begin{table}
\begin{tabulary}{\linewidth}{l L}
  \toprule
  Name & Description \\
  \midrule
  App \texttt{isRunnable} & Says an app can be run. \\
  App \texttt{isInstallable} & Says an app can be installed. \\
  App \texttt{isAnApp} & Tells AppPAL that an app exists. \\
  Policy \texttt{isMetBy(}App\texttt{)} & Says a specific policy is met by an app.  This is used to split policies into smaller components which can be reused and composed. \\
  \texttt{hasPermission(}App, Permission\texttt{)} & Constraint for testing whether an app has been granted a permission. \\
  \texttt{BeforeHourOfDay(}time\texttt{)} & Constraint used to test whether we're before an hour in the day. \\
  \bottomrule
\end{tabulary}
\caption{Typical AppPAL predicates and constraints}
\label{tab:predicates}
\end{table}

% \begin{description}
%   \item[\texttt{App isRunnable}]
%     Used to indicate that an App meets the install policy for the device.
%     Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.
%
%   \item[\texttt{App isAnApp}]
%     Expresses a simple typing relations.
%     AppPAL, and SecPAL, have a safety condition that all variable mentioned in the fact, are also mentioned in the condition body (as is the case for Datalog and other logic languages).
%     These typing relations allow an appeal to ground facts to be made so the safety condition is satisfied.
%     This typically is necessary when an assertion has a constraint but no body conditions: we add the trivial typing statement to the body to ensure the assertion is accepted by AppPAL.
%
%   \item[\texttt{Policy isMetBy(App)}]
%     Policies let you split app behaviour into sub-policies.
%     For example in \autoref{sec:idea} we showed how Emma's installation policy could be written using three \texttt{isMetBy} statements.
%     Splitting policies allows greater control of how each one is checked.
%     We can delegate checking a policy to an expert using the \code{can-say} statement.
%     We can specify more detailed checks using the \code{can-act-as} statements.
%
%   \item[\texttt{Policy shownIsMetBy(Evidence, App)}]
%     A variant of the \texttt{isMetBy} statement that allows some evidence (a proof) to be given that shows the policy is met.
% \end{description}

@@

Earlier we described a \emph{no-tracking-policy} to prevent a user's location being tracked.
In Bob's policy we checked this by checking the permissions of the app.
If the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~Taintdroid~\cite{Fritz:2013vi}).
Our policy now becomes:

@@

% AppPAL also helps attribute blame when things go wrong.
% By modelling the trust relationships in systems we can work out precisely where mistakes were made.
% A recent example demonstrating this is \textsc{CVE-2015-2077}.
% Lenovo was found to be shipping laptops with the \emph{Superfish} malware pre-installed.
% Lenovo had also installed an \ac{av} package on their laptops; this \ac{av} did recognise Superfish as malware.
% Unfortunately they had configured it to ignore Superfish.
% The malware was an ad framework designed to show users products similar to those they viewed on the web.
% Unfortunately it did so by man-in-the-middling all SSL traffic using a shared private key.
%
% From a user's perspective there was a delegation of trust to Lenovo to detect malware.
% Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
% A user might wonder where the breach of trust occurred and who is to blame?
% Is it the \ac{av} failing to spot the malware?
% Has someone else configured the software incorrectly?
% How should they fix the problem?
%
% If we write this in the AppPAL policy language the cause of the breach of trust becomes apparent; as does the fix.
% \begin{lstlisting}[numbers=left, escapeinside={@}{@}, label={lst:lenovo}]
% "user" says "lenovo" can-say inf File isSafe. @\label{lst:lenovo_trust}@
% "lenovo" says "mcafee" can-say inf File isSafe.
% "lenovo" says "C:\System\superfish" isSafe. @\label{lst:lenovo_unsafe}@
% \end{lstlisting}
% The fault lies with \autoref{lst:lenovo_unsafe}.
% Lenovo has caused the problem (they \emph{said} it was safe).
% The fix is to revoke this statement, or to revoke \autoref{lst:lenovo_trust} and find a different \ac{av} supplier.

\subsection{Policy checking}
\label{ssec:eval}

Since AppPAL is an instantiation of SecPAL the policy checking rules are the same.
We do not use Becker~\etal's DatalogC~\cite{Li:2003ix}~based translation and evaluation algorithm, as no DatalogC library exists for Android.
We have implemented the rules directly in Java.
Pseudo-code is given in \autoref{fig:pseudocode}.
Like Becker~\etal we make use of an assertion context to store known statements.
We also store intermediate results to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.

% \begin{figure}
% \[\footnotesize
% % AC, D ⊧ A says B can-act-as C   AC, D ⊧ A says C verbphrase
% % ----------------------------------------------------------- (can-act-as)
% %                 AC, D ⊧ A says B verbphrase
% \infer[\text{\tiny can-act-as}]{
%   AC, D \models \text{\tt A says B verbphrase}}{%
%   AC, D \models \text{\tt A says B can-act-as C} & %
%   AC, D \models \text{\tt A says C verbphrase}}
% \]
% \[\footnotesize
% % AC, ∞ ⊧ A says B can-say D fact   AC, D ⊧ B says fact
% % ----------------------------------------------------- (can-say)
% %                   AC, ∞ ⊧ A says fact
% \infer[\text{\tiny can-say}]{
%   AC, \infty \models \text{\tt A says fact}}{%
%   AC, \infty \models \text{\tt A say B can-say $D$ fact} & %
%   AC, D \models \text{\tt B says fact}}
% \]
% \[\footnotesize
% % (A says fact if fact₁, …, factₖ, c) ∈ AC
% % AC, D ⊧ A says factᵢθ  ∀i ∈ {1…k}    ⊧cθ    vars(factθ) = ∅
% % ----------------------------------------------------------- (cond)
% %                    AC, D ⊧ A says factθ
% \infer[\text{\tiny cond}]{
%   AC, D \models \text{\tt A says fact$\theta$}}{
%   \begin{array}{c}
%     (\text{\tt A says fact if fact$_1$, \ldots, fact$_k$ where c}) \in AC \\
%     \forall i \in 1\cdots k.\;AC, D \models \text{\tt A says fact$_i\theta$}
%   \end{array} &
%   \models \mathtt{c}\theta &
%   \text{\sl vars}(\text{\tt fact$\theta$}) = \emptyset}
% \]
% \caption{AppPAL's evaluation rules.}
% \label{fig:rules}
% \end{figure}

@@

% In SecPAL there is no type.
% All entities are either constants (implemented as strings) or variables.
% There is no differentiation made between a \emph{voiced} constant (who says assertions),
% and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
% In AppPAL the subjects are usually apps or policies.
% Apps do not utter facts;
%   so there is no need to check them as potential speakers of \emph{can-say} statements.
% Similarly if a constant never appears as a subject of a statement, as most speakers do not,
%   there is no need to check it as a possible subject for a \emph{can-act-as} statement.
% We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.
% This means we need to search less constants when attempting to replace a variable, speeding evaluation.

\begin{figure}
\begin{minipage}[b]{0.5\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
def evaluate(ac, rt, q, d)
  return rt[q, d] if rt.contains q, d
  p = cond(ac, rt, q, d)
  return (Proven, rt.update q, d, p) if p.isValid
  p = canSay_CanActAs(ac, rt, q, d)
  return (Proven, rt.update q, d, p) if p.isValid
  return (Failure, rt.update q, d, Failure)

def canSay_CanActAs(ac, rt, q, d)
  ac.constants.each do |c|
    if c.is_a :subject
      p = canActAs ac, rt, q, d
      return Proven if p.isValid
    elsif c.is_a :speaker
      p = canSay ac, rt, q d
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\begin{minipage}[b]{0.5\linewidth}
\begin{lstlisting}[language=Ruby, basicstyle=\ttfamily, keywordstyle=\scriptsize\slshape, columns=flexible]
def cond(ac, rt, q, d)
  ac.add q.fetch if q.isFetchable
  ac.assertions.each do |a|
    if (u = q.unify a.consequent) &&
       (a = u.substitution a).variables == none
      return checkConditions ac, rt, a, d
  return Failure

def checkConditions(ac, rt, a, d)
  getVarSubstitutions(a,ac.constants).each do |s|
    sa = s.substitute a
    if sa.antecedents.all
        { |a| evaluate(ac, rt, a, d).isValid }
      p = evaluateC sa.constraint
      return Proven if p.isValid
  return Failure
\end{lstlisting}
\end{minipage}
\caption{Partial-pseudocode for AppPAL evaluation.}
\label{fig:pseudocode}
\end{figure}
